CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-12-14 codecheck-paper.tex
fix rendering
add two more references and fix some citations

\documentclass[12pt]{article}
\usepackage{f1000_styles}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\usepackage[hyphenbreaks]{breakurl}
\usepackage[hyphens]{url}
\usepackage{setspace}

@@

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Use "execution" insteand of "sharing" in title?
\item Can we introduce "research compendium" as a term?
\item Should we make code and data sharing a prerequisite or leave that door open?
\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
\item Should we ask the people currently listed in the acknowledgements?
\item revisit \url{https://github.com/codecheckers/discussion}
\item 
\end{itemize}

@@

\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
the author, and checks if some/all of the results from the paper can be
reproduced. (See later for exact nature of ``reproduced''). If there are
any problems running the code, the CODECHECKER asks the author for help
to resolve the problems. The burden to provide reproducible material lies with
the author.
The CODECHECKER then tries to run the code again.
This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code does not work (\emph{what do in this
case}). As part of this process, the CODECHECKER could work entirely
locally on their own compute resource, or in the cloud, e.g.~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
are more tailored for scientific publications while others offer commercial
options to, e.g., publishers \cite{}.
Using such cloud-based 
infrastructure allows for the CODECHECKER and author to collaboratively improve
the code and enforces a much more complete definition of the computing environment,
but, unless infrastructure is provided by, e.g., by the publisher, requires code
and data to be published openly online.
Note that the taks of the CODECHECKER is to check the ``mechanics'' of the
workflow, not to evaluate its correctness. 
Stodden et~al. \cite{stodden_setting_2013} destinguish between verification and
validation in the context of Mathematics, and in the same vein a CODECHECK
ensures a verification\footnote{Note this is a different use as the \emph{Verification Reports} article type in the journal \emph{Cortex}
\cite{chambers_verification_2020}},
i.e., checking that the code correctly produces the 
output it claims to create, but not a validation, i.e., does the code implement
the right algorithm to solve the problem that is researched, 
of the computational results.

@@

The topic of blinding is broadly discussed, not the least in the push towards
open peer review as part of the Open Science movement [REF]. Without taking
a strong stand on this discussion, the general motivation for more 
transparency and reproducibility does indeed favour a more open review 
process. However, anonymity can also serve to protect individuals 
\ref{tennant_limitations_2020}, e.g.,  juniour scientists, whereas it might 
be acceptable to require revealing of the names of tenured researchers.
The possible negative effects of a non-anonymous review are greatly reduced
if a CODECHECK is not relevant for the decision to accept or reject, but that
is of course not desirable. Instead, we argue that a CODECHECK is technical
process that should in any case be fixable and not a question of opinion or
faulty approach. It is possible for a nonsense workflow to receive a 
CODECHECK certificate.
Furthermore, the level of documentation that is needed for third parties
to reproduce a workflow is extremely hard to get right. Too often, this 
uncertainty leads to researchers giving up and not documenting at all.

@@

\textbf{4. Workflows must be auditable.}
This requires that the codechecker 
has at least enough material to validate the workflow outputs submitted by 
the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'
and the ICERM report \cite{} defines the level ``Auditable Research'' 
similarly. Note that openness is not required here, as to accommodate 
communities of practice and review workflows that (currently) do not want
to or are not able to work openly.
A completed check eans that code could be executed at least once without
critical errors or
warnings using the provided instructions, and, therefore, all code and data 
was given and could be investigated more deeply or extended in the future.
Ideally, this is a "one~click" step, but achieving this requires particular 
skills. It is also hard to create a sufficient level of documentation for 
third parties. Furthermore automation may lead to people gaming the system
or reliance on technology hiding important details. All such aspects can
reduce the understandability of the material, which we estimate to be the
best mean to ensure long-term transparency and usefulness.
That is why the possibility to communicate is more important
than perfected workflows, and why a CODECHECK is not automated.
We acknowledge that others have argued in favour of bitwise reproducibility
because in the long run it can be automated\footnote{Konrad Hinsen on 
Twitter: \url{https://twitter.com/khinsen/status/1242842759733665799}};
but until then we need CODECHECK's approach.

@@

A number of journals or conferences/workshops provide guidelines for ensuring reproducible publications and their verification.
These include, for example, \ldots{}
% first glimpse into "Code exeuction processes in scholarly publishing" ?
% `https://docs.google.com/document/d/1fMWrFvBTwAYXg5J_bYIjgb-gY7DfQ0DOjx9Vlg_Rbmg/edit# ?

\emph{Information Systems}, with an invited reproducibility paper track
\footnote{\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}}

However, no approach besides CODECHECK spans across journals and is
intentionally designed to be picked up and implemented by many different 
journals or events, and to build a community of codecheckers. CODECHECK shares
this interdisciplinary nature with other community iniatives concered with 
reproducibility awareness, education, and supports, such as ReproHack
\footnote{\url{https://reprohack.github.io/reprohack-hq/}}, Code Copilot
\footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code
\footenote{\url{https://paperswithcode.com/about}}.
The latter recently announced a collaborations with the preprint server
\emph{arXiv} on providing data and code supplements for machine learning
manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}} and runs a reproduciblity challenge
\footnote{\url{https://paperswithcode.com/rc2020}}.

@@

\textbf{Just peer review}: wait, isn't CODECHECK what peer review should be 
doing already? That is correct.
However, the general observation is that the increasing
number of researchers, submissions, and reviews to conduct under the
pressure to publish, puts the scholarly publication system under a lot
of strain [REF].  We also see parallels to the terms 
`reproducible research' and `open science'. Establishing a CODECHECK 
process is largely an acknowledgement that peer reviewing practices 
have faltered to adapt to the challenges of digitisation and computer-based
research or data science. The concept of a CODECHECK, just as the 
reproducibility of research and the openness of science, may be of 
transitional nature. If the activites described here as being part of a 
CODECHECK become the norm for scientific reviews and publications, then
the initiative will have succeeded.
However, the rapidly changing technology used in research leads us to 
believe that a seperate role of codechecker and the expertise it brings
will be benefitial in the long term.

>>>>

\documentclass[12pt]{article}
\usepackage{f1000_styles}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\usepackage{url}
\usepackage{setspace}

@@

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Use "execution" insteand of "sharing" in title?
\item Can we introduce "research compendium" as a term?
\item Should we make code and data sharing a prerequisite or leave that door open?
\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
\item Should we ask the people currently listed in the acknowledgements?
\item revisit \url{https://github.com/codecheckers/discussion}
\end{itemize}

@@

\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
the author, and checks if some/all of the results from the paper can be
reproduced. (See later for exact nature of ``reproduced''). If there are
any problems running the code, the CODECHECKER asks the author for help
to resolve the problems. The burden to provide reproducible material lies with
the author.
The CODECHECKER then tries to run the code again.
This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code does not work (\emph{what do in this
case}). As part of this process, the CODECHECKER could work entirely
locally on their own compute resource, or in the cloud, e.g.~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
are more tailored for scientific publications while others offer commercial
options to, e.g., publishers (cf. \cite{konkol_publishing_2020}).
Using such cloud-based 
infrastructure allows for the CODECHECKER and author to collaboratively improve
the code and enforces a much more complete definition of the computing environment,
but, unless infrastructure is provided by, e.g., by the publisher, requires code
and data to be published openly online.
Note that the taks of the CODECHECKER is to check the ``mechanics'' of the
workflow, not to evaluate its correctness. 
Stodden et~al. \cite{stodden_setting_2013} destinguish between verification and
validation in the context of Mathematics, and in the same vein a CODECHECK
ensures a verification\footnote{Note this is a different use as the \emph{Verification Reports} article type in the journal \emph{Cortex}
\cite{chambers_verification_2020}},
i.e., checking that the code correctly produces the 
output it claims to create, but not a validation, i.e., does the code implement
the right algorithm to solve the problem that is researched, 
of the computational results.

@@

The topic of blinding is broadly discussed, not the least in the push towards
open peer review as part of the Open Science movement 
(cf. \cite{ross-hellauer_guidelines_2019}).
Without taking a strong stand on this discussion, the general motivation 
for more  transparency and reproducibility does indeed favour a more open 
review process.
However, anonymity can also serve to protect individuals 
\ref{tennant_limitations_2020}, e.g.,  juniour scientists, whereas it might 
be acceptable to require revealing of the names of tenured researchers.
The possible negative effects of a non-anonymous review are greatly reduced
if a CODECHECK is not relevant for the decision to accept or reject, but that
is of course not desirable. Instead, we argue that a CODECHECK is technical
process that should in any case be fixable and not a question of opinion or
faulty approach. It is possible for a nonsense workflow to receive a 
CODECHECK certificate.
Furthermore, the level of documentation that is needed for third parties
to reproduce a workflow is extremely hard to get right. Too often, this 
uncertainty leads to researchers giving up and not documenting at all.

@@

\textbf{4. Workflows must be auditable.}
This requires that the codechecker 
has at least enough material to validate the workflow outputs submitted by 
the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'
and the ICERM report \cite{stodden_setting_2013} defines the level
``Auditable Research''  similarly.
Note that openness is not required here, as to accommodate 
communities of practice and review workflows that (currently) do not want
to or are not able to work openly.
A completed check eans that code could be executed at least once without
critical errors or
warnings using the provided instructions, and, therefore, all code and data 
was given and could be investigated more deeply or extended in the future.
Ideally, this is a "one~click" step, but achieving this requires particular 
skills. It is also hard to create a sufficient level of documentation for 
third parties. Furthermore automation may lead to people gaming the system
or reliance on technology hiding important details. All such aspects can
reduce the understandability of the material, which we estimate to be the
best mean to ensure long-term transparency and usefulness.
That is why the possibility to communicate is more important
than perfected workflows, and why a CODECHECK is not automated.
We acknowledge that others have argued in favour of bitwise reproducibility
because in the long run it can be automated\footnote{Konrad Hinsen on 
Twitter: \url{https://twitter.com/khinsen/status/1242842759733665799}};
but until then we need CODECHECK's approach.

@@

A number of journals or conferences/workshops provide guidelines for ensuring reproducible publications and their verification.
These include, for example, \ldots{}
% first glimpse into "Code exeuction processes in scholarly publishing" ?
% `https://docs.google.com/document/d/1fMWrFvBTwAYXg5J_bYIjgb-gY7DfQ0DOjx9Vlg_Rbmg/edit# ?
\emph{Information Systems}, with an invited reproducibility paper track
\footnote{\url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}}

Another journal initiative is more indirectly connected with reproductions.
\emph{Nature Machine Intelligence} recently introduced a new type of article,
the reusability report \cite{noauthor_research_2020}.
Inspired by the detailed and nuanced submissions to a reproducibility 
challenge, the reusability report focuses on the exploration of robustness
and generalizability of the original paper's claims
\cite{noauthor_research_2020}. This answers the specific community's 
challenges around computational reproducibility and also values these kind
of contributions as independent publications, which goes beyond the goals 
of CODECHECK.

However, no approach besides CODECHECK spans across journals and is
intentionally designed to be picked up and implemented by many different 
journals or events, and to build a community of codecheckers. CODECHECK shares
this interdisciplinary nature with other community iniatives concered with 
reproducibility awareness, education, and supports, such as ReproHack
\footnote{\url{https://reprohack.github.io/reprohack-hq/}}, Code Copilot
\footnote{\url{https://twitter.com/Code_Copilot}}, or Papers with Code
\footenote{\url{https://paperswithcode.com/about}}.
The latter recently announced a collaborations with the preprint server
\emph{arXiv} on providing data and code supplements for machine learning
manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}} and runs a reproduciblity challenge
\footnote{\url{https://paperswithcode.com/rc2020}}.
% Papers with Code now also does more disciplines than just ML.
Likewise, different disciplines and journals provide reproducibility
checklists, e.g., \cite{rosenberg_next_2020}, which naturally share some
aspects while addressing particularities as well as addressing researchers
from different fields. Regarding the education and guidance for authors, 
we see CODECHECK's role in referencing and linking educational efforts
and helpful material, not as a creator and maintainer of content.

@@

\textbf{Just peer review}: wait, isn't CODECHECK what peer review should be 
doing already? That is correct.
However, the general observation is that the increasing
number of researchers, submissions, and reviews to conduct under the
pressure to publish, puts the scholarly publication system under a lot
of strain.  We also see parallels to the terms 
`reproducible research' and `open science'. Establishing a CODECHECK 
process is largely an acknowledgement that peer reviewing practices 
have faltered to adapt to the challenges of digitisation and computer-based
research or data science. The concept of a CODECHECK, just as the 
reproducibility of research and the openness of science, may be of 
transitional nature. If the activites described here as being part of a 
CODECHECK become the norm for scientific reviews and publications, then
the initiative will have succeeded.
However, the rapidly changing technology used in research leads us to 
believe that a seperate role of codechecker and the expertise it brings
will be benefitial in the long term.
